---
title: "Test All Cycles"
author: "Matthew Hofkes"
date: "2023-10-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


This code tests all 3305 cycles of the Denver Water headloss data for anomalies.  Cycles under 60 measurements (5 hours) are considered "short cycles" and not tested.  The code implements 
1. the hybrid (elbow) approach 
2. the hybrid (AICc) approach
3. the Bayesian approach
4. EWMA
5. CUSUM 
and then produces confusion tables for the results of each.  The code should be run in order.


```{r}
load("Separated_Filter_Cycles/headloss_matrix.rda")
load("Separated_Filter_Cycles/headloss_cycles_info.rda")
source("hybrid.smoother.freq.R")
source("hybrid.smoother.bayes.R")
```


Frequentist (Elbow)
```{r}
all.analyzed.cycles.list.freq <- vector(mode = "list", length = 3305)
all.cycles.anomaly.table.freq <- data.frame(cycle.number=1:3305,cycle.length=headloss_cycles_info$cycle_length-6,Anomalous=NA,Identified=NA,RMSE=NA,Max.CP=NA,run.time=NA)

for (cn in 1:3305){
  if (headloss_cycles_info$cycle_length[cn] > 60){
    n <- headloss_cycles_info$cycle_length[cn]+1
    x <- 1:(n-6)
    y <- headloss_matrix[cn,4:(n-3)]
    n <- length(x)
  
    lambda.grid <- 2^seq(2,-6,-1)
    omega.grid <- 2^seq(4,16,1)
    
    init.time <- proc.time()
    model <- hybrid.smoother.freq(x,y,lambda.grid,omega.grid,tolerance = 1e-4,method="elbow")
    time <- time <- (proc.time()-init.time)[3]
    all.analyzed.cycles.list.freq[[cn]] <- model
    names(all.analyzed.cycles.list.freq)[cn] <- paste0("model.",cn)
    all.cycles.anomaly.table.freq[cn,4] <- ifelse(length(model$change.points)!=0,"YES","NO")
    all.cycles.anomaly.table.freq[cn,5] <- sqrt(sum(model$residuals^2)/n)
    all.cycles.anomaly.table.freq[cn,6] <- ifelse(length(model$change.points)!=0, max(abs(model$gamma.coefs)[model$change.points]),NA)
    all.cycles.anomaly.table.freq[cn,7] <- time
  }
  
  if (headloss_cycles_info$cycle_length[cn] < 60){
    all.analyzed.cycles.list.freq[[cn]] <- "Short Cycle"
  }
  
  if(cn%%50==0) cat(cn," ")
  
}

```


Frequentist (AICc)
```{r}
all.analyzed.cycles.list.freq.AIC <- vector(mode = "list", length = 3305)
all.cycles.anomaly.table.freq.AIC <- data.frame(cycle.number=1:3305,cycle.length=headloss_cycles_info$cycle_length-6,Anomalous=NA,Identified=NA,RMSE=NA,Max.CP=NA,run.time=NA)

for (cn in 1:3305){
  if (headloss_cycles_info$cycle_length[cn] > 60){
    n <- headloss_cycles_info$cycle_length[cn]+1
    x <- 1:(n-6)
    y <- headloss_matrix[cn,4:(n-3)]
    n <- length(x)
  
    lambda.grid <- 2^seq(2,-4,-1)
    omega.grid <- 2^seq(4,16,1)
    
    init.time <- proc.time()
    model <- hybrid.smoother.freq(x,y,lambda.grid,omega.grid,tolerance = 1e-4,method="AICc")
    time <- time <- (proc.time()-init.time)[3]
    all.analyzed.cycles.list.freq.AIC[[cn]] <- model
    names(all.analyzed.cycles.list.freq.AIC)[cn] <- paste0("model.",cn)
    all.cycles.anomaly.table.freq.AIC[cn,4] <- ifelse(length(model$change.points)!=0,"YES","NO")
    all.cycles.anomaly.table.freq.AIC[cn,5] <- sqrt(sum(model$residuals^2)/n)
    all.cycles.anomaly.table.freq.AIC[cn,6] <- ifelse(length(model$change.points)!=0, max(abs(model$gamma.coefs)[model$change.points]),NA)
    all.cycles.anomaly.table.freq.AIC[cn,7] <- time
  }
  
  if (headloss_cycles_info$cycle_length[cn] < 60){
    all.analyzed.cycles.list.freq.AIC[[cn]] <- "Short Cycle"
  }
  
  if(cn%%50==0) cat(cn," ")
  
}

```



Bayesian Approach
```{r}
all.analyzed.cycles.list.bayes <- vector(mode = "list", length = 3305)
all.cycles.anomaly.table.bayes <- data.frame(cycle.number=1:3305,cycle.length=headloss_cycles_info$cycle_length-6,Anomalous=NA,Identified=NA,s=NA,max.gamma=NA,run.time=NA)

for (cn in 1:3305){
  if (headloss_cycles_info$cycle_length[cn] > 60){
    n <- headloss_cycles_info$cycle_length[cn]+1
    x <- 1:(n-6)
    y <- headloss_matrix[cn,4:(n-3)]
    n <- length(x)
  
    init.time <- proc.time()
    model <- hybrid.smoother.bayes(x,y,n.chains = 2)
    time <- time <- (proc.time()-init.time)[3]
    
    all.analyzed.cycles.list.bayes[[cn]] <- model
    names(all.analyzed.cycles.list.bayes)[cn] <- paste0("model.",cn)
    all.cycles.anomaly.table.bayes[cn,4] <- ifelse(length(model$change.points)!=0,"YES","NO")
    all.cycles.anomaly.table.bayes[cn,5] <- model$s
    all.cycles.anomaly.table.bayes[cn,6] <- ifelse(length(model$change.points)!=0, max(abs(model$gammas)),0)
    all.cycles.anomaly.table.bayes[cn,7] <- time
  }
  
  if (headloss_cycles_info$cycle_length[cn] < 60){
    all.analyzed.cycles.list.bayes[[cn]] <- "Short Cycle"
  }
  
  if(cn%%20==0) cat(cn," ")
  
}
```




CUSUM
```{r}
library(qcc)
cusum.with.reset <- function(data,center=0,decision.interval=4,se.shift=.5){
  data <- data - center
  x.n <- length(data)
  # sd.dd <- sd(data)
  sd.dd <- sd.xbar.one(data)
  # sd.dd <- cusum(detrended.data,decision.interval = 4,se.shift=.5)$std.dev
  slack <- se.shift
  limit <- decision.interval
  cs.pos <- 0
  cs.neg <- 0
  cs.pos.vec <- rep(0,x.n)
  cs.neg.vec <- rep(0,x.n)
  for ( i in 1:x.n){
    cs.pos <- max(cs.pos + data[i]/sd.dd - slack,0)
    cs.neg <- min(cs.neg + data[i]/sd.dd + slack,0)
    cs.pos.vec[i] <- cs.pos
    cs.neg.vec[i] <- cs.neg
    
    if (cs.pos > limit | cs.neg < -limit){
      cs.pos <- 0
      cs.neg <- 0
    }
    
    violations <- sort(c(which(cs.pos.vec>limit),which(cs.neg.vec< -limit)))
  }
  violations
}

all.analyzed.cycles.list.cusum <- vector(mode = "list", length = 3305)
all.cycles.anomaly.table.cusum <- data.frame(cycle.number=1:3305,cycle.length=headloss_cycles_info$cycle_length-6,Anomalous=NA,Identified=NA,s=NA,max.gamma=NA,run.time=NA)

for (cn in 1:3305){
  if (headloss_cycles_info$cycle_length[cn] > 60){
    n <- headloss_cycles_info$cycle_length[cn]+1
    x <- 1:(n-6)
    y <- headloss_matrix[cn,4:(n-3)]
    n <- length(x)
  
    init.time <- proc.time()
    ### Detrend
    detrended.data <- Tps(x,y)$residuals
    ### CUSUM
    violations <- cusum.with.reset(detrended.data,center=0,decision.interval=4,se.shift=.5)
    time <- time <- (proc.time()-init.time)[3]
    
    all.analyzed.cycles.list.cusum[[cn]] <- violations
    names(all.analyzed.cycles.list.cusum)[cn] <- paste0("model.",cn)
    all.cycles.anomaly.table.cusum[cn,4] <- ifelse(length(violations)!=0,"YES","NO")
    # all.cycles.anomaly.table.cusum[cn,5] <- model$s
    # all.cycles.anomaly.table.cusum[cn,6] <- ifelse(length(violations)!=0, max(violations),0)
    all.cycles.anomaly.table.cusum[cn,7] <- time
  }
  
  if (headloss_cycles_info$cycle_length[cn] < 60){
    all.analyzed.cycles.list.cusum[[cn]] <- "Short Cycle"
  }
  
  if(cn%%20==0) cat(cn," ")
  
}
```




EWMA
```{r}
library(qcc)

all.analyzed.cycles.list.ewma <- vector(mode = "list", length = 3305)
all.cycles.anomaly.table.ewma <- data.frame(cycle.number=1:3305,cycle.length=headloss_cycles_info$cycle_length-6,Anomalous=NA,Identified=NA,s=NA,max.gamma=NA,run.time=NA)

for (cn in 1:3305){
  if (headloss_cycles_info$cycle_length[cn] > 60){
    n <- headloss_cycles_info$cycle_length[cn]+1
    x <- 1:(n-6)
    y <- headloss_matrix[cn,4:(n-3)]
    n <- length(x)
  
    init.time <- proc.time()
    ### Detrend
    detrended.data <- Tps(x,y)$residuals
    ### EWMA
    model <- ewma(detrended.data,center=0,plot=FALSE,lambda=.3)
    violations <- unique(c(unique(model$violations), unique(model$violations) + 1, unique(model$violations) - 1)) 
    time <- time <- (proc.time()-init.time)[3]
    
    all.analyzed.cycles.list.ewma[[cn]] <- violations
    names(all.analyzed.cycles.list.ewma)[cn] <- paste0("model.",cn)
    all.cycles.anomaly.table.ewma[cn,4] <- ifelse(length(violations)!=0,"YES","NO")
    # all.cycles.anomaly.table.ewma[cn,5] <- model$s
    # all.cycles.anomaly.table.ewma[cn,6] <- ifelse(length(violations)!=0, max(violations),0)
    all.cycles.anomaly.table.ewma[cn,7] <- time
  }
  
  if (headloss_cycles_info$cycle_length[cn] < 60){
    all.analyzed.cycles.list.ewma[[cn]] <- "Short Cycle"
  }
  
  if(cn%%20==0) cat(cn," ")
  
}
```








Locations of anomalous cycles and completing anomaly tables
```{r}
anomalies <- c(27,38,76,135,136,170,203,205,227,346,369,383,387,434,446,502,615,657,860,927,1077,1309,1339,1374,1400,1421,1471,1536,1592,1594,1659,1670,1679,1704,1836,1841,1899,1945,2009,2018,2038,2094,2182,2225,2269,2320,2377,2409,2424,2433,2515,2540,2633,2655,2724,2760,2861,2885,2905,2963,2994,3019,3114,3190,3201)

all.cycles.anomaly.table.bayes$Anomalous[!is.na(all.cycles.anomaly.table.bayes$Identified)]<-"NO"
for (i in anomalies){
  all.cycles.anomaly.table.bayes$Anomalous[i] <- "YES"
}

all.cycles.anomaly.table.freq$Anomalous[!is.na(all.cycles.anomaly.table.freq$Identified)]<-"NO"
for (i in anomalies){
  all.cycles.anomaly.table.freq$Anomalous[i] <- "YES"
}
  
all.cycles.anomaly.table.freq.AIC$Anomalous[!is.na(all.cycles.anomaly.table.freq.AIC$Identified)]<-"NO"
for (i in anomalies){
  all.cycles.anomaly.table.freq.AIC$Anomalous[i] <- "YES"
}

all.cycles.anomaly.table.cusum$Anomalous[!is.na(all.cycles.anomaly.table.cusum$Identified)]<-"NO"
for (i in anomalies){
  all.cycles.anomaly.table.cusum$Anomalous[i] <- "YES"
}

all.cycles.anomaly.table.ewma$Anomalous[!is.na(all.cycles.anomaly.table.ewma$Identified)]<-"NO"
for (i in anomalies){
  all.cycles.anomaly.table.ewma$Anomalous[i] <- "YES"
}

```



Confusion Tables 
```{r}
library(caret)
```



FREQ (Elbow) Confusion Table

```{r}
#Frequentist (elbow) - With q1 sds Buffer and requirement that disturbance is at least q2 units
q1 <- 0
q2 <- 0
temp.table <- all.cycles.anomaly.table.freq
temp.table$Identified[temp.table$Max.CP<q1*temp.table$RMSE | temp.table$Max.CP<q2] <- "NO" 
predicted <- as.factor(temp.table$Identified[!is.na(temp.table$Identified)])    
expected <- as.factor(temp.table$Anomalous[!is.na(temp.table$Identified)])
confusionMatrix(predicted,expected)
```



Freq (AICc) Confusion Table

```{r}
#Frequentist(AICc) - With q1 sds Buffer and requirement that disturbance is at least q2 units
q1 <- 0
q2 <- .15
temp.table <- all.cycles.anomaly.table.freq.AIC
temp.table$Identified[temp.table$Max.CP<q1*temp.table$RMSE | temp.table$Max.CP<q2] <- "NO" 
predicted <- as.factor(temp.table$Identified[!is.na(temp.table$Identified)])    
expected <- as.factor(temp.table$Anomalous[!is.na(temp.table$Identified)])
confusionMatrix(predicted,expected)
```



Bayes Confusion Table

```{r}
#Bayesian - With no buffer and requirement that disturbance is at least q2 units
q1 <- 0
q2 <- .15
temp.table <- all.cycles.anomaly.table.bayes
temp.table$Identified[temp.table$max.gamma<q1*temp.table$s | temp.table$max.gamma<q2] <- "NO" 
predicted <- as.factor(temp.table$Identified[!is.na(temp.table$Identified)])    
expected <- as.factor(temp.table$Anomalous[!is.na(temp.table$Identified)])
confusionMatrix(predicted,expected)
```


Cusum Confusion Table

```{r}
#Cusum
temp.table <- all.cycles.anomaly.table.cusum
predicted <- as.factor(temp.table$Identified[!is.na(temp.table$Identified)])    
expected <- as.factor(temp.table$Anomalous[!is.na(temp.table$Identified)])
confusionMatrix(predicted,expected)
```

EWMA Confusion Table

```{r}
#EWMA
temp.table <- all.cycles.anomaly.table.ewma
predicted <- as.factor(temp.table$Identified[!is.na(temp.table$Identified)])    
expected <- as.factor(temp.table$Anomalous[!is.na(temp.table$Identified)])
confusionMatrix(predicted,expected)
```



Create Timing Plot
```{r}
# smaller.table.freq <- all.cycles.anomaly.table.freq[!is.na(all.cycles.anomaly.table.freq$Anomalous),]
# smaller.table.bayes <- all.cycles.anomaly.table.bayes[!is.na(all.cycles.anomaly.table.bayes$Anomalous),]
# smaller.table.freq <- smaller.table.freq[!(smaller.table.bayes$run.time>80 & smaller.table.bayes$cycle.length<450),]
# smaller.table.bayes <- smaller.table.bayes[!(smaller.table.bayes$run.time>80 & smaller.table.bayes$cycle.length<450),]
# 
# plot(smaller.table.bayes$cycle.length,smaller.table.bayes$run.time/4,cex=.6,log="xy", main="Runtime Comparison",ylab = "Computation Time (Seconds)",xlab="Cycle Length",col="blue",ylim=c(.05,40),pch=17)
# points(smaller.table.freq$cycle.length,smaller.table.freq$run.time*1/3,log="xy",cex=.6,col="red",pch=16)
# legend(60, 25, c("Frequentist (Elbow)","Bayesian"), col=c("red","blue"),pch=c(16,17))
```


